How to Analyze Build Logs --- A~Comparative Study of Chunk~Retrieval~Techniques

### 2020-04-23 tse-paper/content.tex
passives

To answer this question, in the second step of the study
(\Cref{fig:overview}), we build and compare three
implementations that each stand protoypical for one of the idenitied
strategies to analyze build logs:

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
% TODO maybe make the point here that it is questionable whether this
%includes a comprehensive tokenizer, lexer etc.
%based solution and that parsing in many cases really meant regexp.
If we could figure that out, then we would assign it to regexp instead.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with manual categorizing build logs and selecting
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.

@@

\subsubsection{Which build logs are analyzed? (RQ 1.3)}
\Cref{fig:litsur:log_producer} shows the distribution of supported
build tools.
Of the articles we surveyed, 30\% analyzed build logs produced by
Maven~\cite{maven2019website}, 16\% analyzed logs produced by
Gradle~\cite{gradle2020website}
and 10\% analyzed logs produced by Ant~\cite{ant2020website}.

@@

In summary, the results of the literature study show the need for
build log
analysis techniques that are automated, build-environment agnostic
% TODO who is the user?
and can be configured by the user with little effort.
% We achieve this by developing three techniques for which the
% information to be retrieved is specified through examples or keywords.

% TODO: the following paragraphs have nothing to do with a discussion
%of the literature survey results.
% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
Examples consist of logs and corresponding chunks.
The examples and keywords are provided by the user and can simply be
exchanged whenever new cases are introduced.

@@

Our evaluation of the three chunk retrieval techniques is done on the
\emph{LogChunks} data set because it encompasses a broad range of
build environments which enables us to measure the generic
applicability of the techniques.
The targeted chunks are describing error messages or
``the reason the build failed,'' which was also the information
most targeted by the articles in our survey.

@@

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
The
granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means executing a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
Our implementation is based on the
\emph{PROSE library}~\cite{prose2019webpage}.
The PROSE library is powered by the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that should be retrieved by the synthesized program when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary would be too complicated to synthesize, PBE returns the
error message ``no program found.''

% TODO Moritz: you said: The following is very obvious and could apply
% to any
% % technique.
% % kinda.
% It does give a descriptive error message on no match (and
% % that is not possible in other techniques)
% % otherwise yes, nothing interesting happens in the application of pbe
% % though for consistency we might need this section? or merge
% config/application?
% TODO Why does it return an error message? Shouldn't it also return
% <empty string>? Does it not rather return an error message if no
% program can be synthesized?
% TODO Moritz: a regular expression can match an empty string
% and that is different from not finding a match
% it returns error messages in both cases
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program or an error message if the
program found no match.

@@

To fully comprehend how difficult a task build log analysis really
is without any a-priori knowledge, we include a comparison with a
base line generated by Random Line Retrieval (RLR), that is,
picking lines randomly from the build log.
RLR mimics the
situation of guessing blindly which lines might be interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples, thus arguably
giving it a slight advantage.

@@

When PBE is trained with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if the third example from \Cref{lst:chunk-example-3},
which has a
different structural representation, is added to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

KWS is configured with keywords that appear close to the chunk within
the build log.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE,'' ``wrong,'' or ``failed''.
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL.''
If KWS was to be trained with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

\subsection{Tool Implementation}
For the comparison study, we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is implemented in Ruby and
calls the separate technique implementations over the command line.
We implemented PBE in C\# based on the Microsoft PROSE
library~\cite{prose2019webpage} and CTS, KWS, and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.

@@

% TODO Say something about this being a pecularity of PBE ...
\Cref{fig:failure-reason-PBE} shows the results of chunk
retrieval with PBE.
Out of the overall 400 runs, 5 per each one of the 80 example
sets, PBE extracted all desired lines successfully in 138 cases.
In 59 cases, a regular expression program could be synthesized, though
it did not find match on the test build log.
In 30 cases, the synthesized program only extracted a
subset of the desired lines.
For these 30 cases, the average recall
was 28\%.
% TODO: Perhaps we should call it partially successful? Let's
% discuss this in a meeting
\Cref{lst:pbe-unsuccessful} shows how an example of such
an \emph{unsuccessful retrieval} looks like.
In it, the synthesized program only
retrieved two of the four targeted lines.
In 173 of the 400 cases
could PROSE not synthesize a program consistent with all of the training
examples.

@@

\Cref{fig:failure-reason-categorycount-PBE} shows the results of PBE
runs depend on the number of structural categories in the training and
test examples.
The figure demonstrates that program synthesis mostly
returns exactly the desired output when there are is only one
structural category present in the training and test examples.
However, when
two or more structural categories are simultaneously present,
PROSE could in most cases not synthesize a program.
For four or more present
categories PROSE could never synthesize a consistent program.

@@

\Cref{fig:recall-precision-categorycount-KWS} shows the same
measurements for an increasing number of structural categories in the
training and test examples.
For more than one structural category in
the training and test examples the recall decreases by about 20\% and
the precision decreases about 6\%.
For more than two structural
categories no clear trend is visible in precision, recall, or
F$_{1}$-score.

@@

\noindent
\textbf{Configuration and Input.}
Our study results show that chunk retrieval with PBE gives best
results when the log chunks are structurally identical.
This is
because PROSE has difficulty synthesizing OR-based
programs~\cite{mayer2015user}.
PBE is thus suited to retrieve information
chunks that always have the same defining surrounding or internal
structure.
To extract for example the reason a build failed, the log
passage describing the failure would always have to be started and
ended the same way.

@@

% shouldn't it also better be in threats to validity?
% it is as well.
back for my thesis you said we should put it
% directly here to make it more prominent avoid angry critics
% I think it's good here
% but it's also self-contained so feel free to comment it out
\textbf{Caveat!}
This is a preliminary hypothesis based on the results
from our comparison study.
The recommendations could therefore be
influenced by idiosyncrasies of our specific implementation of the
chunk retrieval techniques as well as the \emph{LogChunks}
data set.

@@

\textbf{Implementation.}
Our results depend on the implementation of the investigated chunk
retrieval techniques and the libraries we used.
Our implementation of
PBE is based on the program synthesis provided by PROSE\@.
The
idiosyncrasies of this framework influence the PBE results.
Other
frameworks similar to PROSE might have somewhat different strengths
and weaknesses.
For example, the need for examples from a single
structural category stems from the fact that PROSE cannot learn
regular expression programs with arbitrary boolean conjunctions and
disjunctions~\cite{mayer2015user}.
PROSE introduced this constraint to
keep the synthesis performance reasonable.
At the same time, this is
clearly a current theoretical challenge of all PBE implementations,
and we therefore attribute it to the technique itself, rather than the
specific implementation.

Our implementation of CTS is dependent on the library
{\tt text2vec}~\cite{text2vec2019webpage}
and the way it splits strings into word tokens.
On the other hand,
{\tt text2vec} is the de-facto standard library to do word embeddings
in R.
We intentionally chose a simple, minimally configured or tuned
approach to compare against.
Tuning the text similarity
meta-parameters more to the specific use case of chunk retrieval from
build logs would yield better chunk retrieval results.

@@

\textbf{Data Set.}
The outcomes of the comparison study are highly dependent on the build
logs from the \emph{LogChunks} data set.
It consists of build
logs from open source projects and therefore it is not clear whether
our results apply to industry projects that use build tools
not popular in open source development.
However, since \emph{LogChunks} covers a wide array of build tools and
programming languages,
we believe that the findings do generalize.
We collected build logs from Travis CI.
Yet, the format of the log chunk we chose
for the evaluation is largely independent of Travis CI\@.
This
is because the reason the build failed is described within the build
logs by the tools themselves and not the Travis CI environment.

\textbf{Training Set Size.}
Especially the results for CTS might be influenced by the fact that we
only trained on one to five examples.
We chose this small training set
size as the training examples have to be provided per repository and
we expect a developer to not want to provide more examples than the
small numbers we evaluated on.
The fact that PBE and CTS performed best with already two training
examples shows that the training set size is large enough to
compare the three techniques we presented.

@@

\textbf{Usability Analysis of Chunk Retrieval Output.}
Our
analysis of the output produced by the chunk retrieval focuses on
precision and recall.
A next step could be to investigate how useful these
outputs really are to developers through controlled experiments or
user studies.

>>>>

To answer this question, in the second step of the study
(\Cref{fig:overview}), we build and compare three
implementations that each stand prototypical for one of the identified
strategies to analyze build logs:

@@

The most mentioned method (34\%) was using a parser, where we also
included
rather imprecise statements like ``we parse the build logs''
~\cite{rahman2018impact}.
% TODO maybe make the point here that it is questionable whether this
%includes a comprehensive tokenizer, lexer etc.
%based solution and that parsing in many cases really meant regexp.
%If we could figure that out, then we would assign it to regexp instead.
25\% of the articles described the use of regular expressions and 20\%
inspected the logs manually.
Another 8\% employed machine learning such as natural language
processing or information retrieval techniques.
For example,
Seo et al.~\cite{seo2014programmers} developed a custom
parser to classify error messages, while Vassallo et
al.~\cite{vassallo2017a-tale} analyzed build logs with regular
expressions.
Ghaleb et al.~\cite{ghaleb2019studying} used a compound approach.
They started with manual categorizing build logs and selecting
search strings that identify the target category.
Based on the search strings they created a script that automatically
classifies the remaining logs.

@@

\subsubsection{Which build logs are analyzed? (RQ 1.3)}
\Cref{fig:litsur:log_producer} shows the distribution of supported
build tools.
Of the articles we surveyed, 30\% analyzed
Maven~\cite{maven2019website} build logs,
16\% analyzed Gradle~\cite{gradle2020website} logs
and 10\% analyzed Ant~\cite{ant2020website} logs.

@@

In summary, the results of the literature study show the need for
build log
analysis techniques that are automated, build-environment agnostic
% TODO who is the user?
and can be configured by a user with little effort.
% We achieve this by developing three techniques for which the
% information to be retrieved is specified through examples or keywords.

% TODO: the following paragraphs have nothing to do with a discussion
%of the literature survey results.
% what we need is some derivation of the three techniques 1) regex 2)
%cts 3) kws, best repeated here in the discussion and then from the lit
%survey above, section 2.4.2
For this reason, we are focussing on techniques that run automatically
after being configured through examples or keywords.
Examples consist of logs and corresponding chunks.
The user provides these examples and keywords, which can simply be
exchanged whenever new cases are introduced.

@@

We evaluate the three chunk retrieval techniques on the
\emph{LogChunks} data set because it encompasses a broad range of
build environments which enables us to measure the generic
applicability of the techniques.
The targeted chunks are describing error messages or
``the reason the build failed,'' which was also the information
most targeted by the articles in our survey.

@@

With the term \textit{configuration}, we abstract over the training
and parametrization that different techniques require in different
forms.
A configuration can be explicitly stated or implicitly derived
by learning through provided examples.
It is therefore a manual
specification of which information the chunk retrieval should target.
It also supplies the necessary information for the technique to
identify the targeted information chunk in a build log.
Each chunk
retrieval technique has a specific \textit{granularity}, i.e.,\ the
smallest piece of text it can return (e.g., a line, or a word).
% The
% granularity might be adjustable by configuration.
\emph{Running a
chunk retrieval technique} means executing a fully configured
technique to consumes as input a build log in plain text format and to
produce an array as output.
The array consists of substrings of the
build log text.

\subsection{Program Synthesis by Example (PBE)}
The concept of \emph{Programming by Example} aims to capture the
intent of the user through examples which they provide.
Leveraging this approach, we implemented a chunk retrieval technique
for build logs.
The \emph{PROSE library}~\cite{prose2019webpage} is the basis of
our implementation.
This library builds on the generic program synthesis framework
\emph{FlashMeta}~\cite{polozov2015flashmeta:} and the specialized
text extraction DSL \emph{FlashExtract}~\cite{le2014flashextract:}.
Both enable us to synthesize regular expression programs
consistent~\cite{mitchell1982generalization} with a set of in-
and output examples given by the user.
The usage of example enables the user to configure a chunk
retrieval without understanding the whole structure of
the build log.

\subsubsection{Configuration}
In- and output example pairs are the main driver of Programming by
Example; we refer to them in short as \emph{examples}.
The \emph{input} is the text of the build log file.
The \emph{output} is
a substring of the log file text, representing the
substring that the synthesized program should retrieve when
given the corresponding input file.
One or multiple examples, the
training set, \emph{configure} a specific chunk retrieval with PBE:
they define the substring of a build log that should be extracted.
The PROSE program synthesis then tries to construct a program
consistent with all training examples.
If PROSE cannot synthesize a program, e.g.,\
because the regular expression
necessary is too complicated to synthesize, PBE returns the
error message ``no program found.''

% Moritz: The following is very obvious and could apply
% to any
% % technique.
% % kinda.
% It does give a descriptive error message on no match (and
% % that is not possible in other techniques)
% % otherwise yes, nothing interesting happens in the application of pbe
% % though for consistency we might need this section? or merge
% config/application?
% Why does it return an error message? Shouldn't it also return
% <empty string>? Does it not rather return an error message if no
% program can be synthesized?
% TODO Moritz: a regular expression can match an empty string
% and that is different from not finding a match
% it returns error messages in both cases
\subsubsection{Application}
A run of PBE takes a build log file as input and applies the
synthesized regular expression program.
It then returns the substring
of the build log matched by the program or an error message if the
program found no match.

@@

To fully comprehend how difficult a task build log analysis really
is without any a-priori knowledge, we include a comparison with a
base line of
picking lines randomly from the build log (RLR).
RLR mimics the
situation of guessing blindly which lines are interesting.
The only configuration option for RLR is the number of lines it should
retrieve.
For a fair comparison to the other techniques (whose number of
returned lines is dynamic), we configure RLR to return the average
number of lines in the chunks of the training examples, thus arguably
giving it a slight advantage.

@@

When we train PBE with these two examples, it produces
a regular expression program similar to the one presented in
\Cref{lst:prose-program-simplified}.
The retrieval
starts after a colon and a line separator and before a capital
letter.
It ends before two line separators,
which is consistent with the two example
chunks from \Cref{lst:chunk-example}.
This shows, that the characters structuring the text around the chunk are
crucial for regular expressions to identify a log chunk.
In fact, if we add the third example from \Cref{lst:chunk-example-3},
which has a
different structural representation, to the training set,
PBE is not able to synthesize a program and returns
``\texttt{no program found}.''
We later express this difference in structural representation through
dividing log chunks into \emph{structural categories}.
The chunks from \Cref{lst:chunk-example}
are in the \emph{same structural category}, while the chunk from
\Cref{lst:chunk-example-3} is in a
\emph{different structural category}.
The results of the empirical comparison study on the chunk retrieval
techniques
show that the structural representation of the targeted chunk greatly
influences
the performance of the chunk retrieval techniques.

@@

Keywords that appear close to the chunk within
the build log configure the technique KWS\@.
For the examples from \Cref{lst:chunk-example} these could be
``FAILURE,'' ``wrong,'' or ``failed''.
The example from \Cref{lst:chunk-example-3} could be found
by searching for ``FAIL.''
If we train KWS with all three examples and the aforementioned
associated keywords, it would search for the three keywords from
\Cref{lst:chunk-example}, as they appear twice---most often within
the training set of three examples.

\subsection{Tool Implementation}
For the comparison study, we implemented each of the chunk retrieval
techniques and a unifying interface.
The unified interface is written in Ruby and
calls the separate technique implementations over the command line.
We implemented PBE in C\# based on the Microsoft PROSE
library~\cite{prose2019webpage} and CTS, KWS, and RLR using
R and the text2vec library~\cite{text2vec2019webpage}.

@@

% TODO Say something about this being a pecularity of PBE ...
\Cref{fig:failure-reason-PBE} shows the results of chunk
retrieval with PBE.
Out of the overall 400 runs, 5 per each one of the 80 example
sets, PBE extracted all desired lines successfully in 138 cases.
In 59 cases, PBE could synthesize a regular expression program, though
it did not find match on the test build log.
In 30 cases, the synthesized program only extracted a
subset of the desired lines.
For these 30 cases, the average recall
was 28\%.
% TODO: Perhaps we should call it partially successful? Let's
% discuss this in a meeting
\Cref{lst:pbe-unsuccessful} shows how an example of such
an \emph{unsuccessful retrieval} looks like.
In it, the synthesized program only
retrieved two of the four targeted lines.
In 173 of the 400 cases
could PROSE not synthesize a program consistent with all of the training
examples.

@@

\Cref{fig:failure-reason-categorycount-PBE} shows the results of PBE
runs depend on the number of structural categories in the training and
test examples.
The figure demonstrates that program synthesis mostly
returns exactly the desired output when there is only one
structural category present in the training and test examples.
However, when
two or more structural categories are simultaneously present,
PROSE could in most cases not synthesize a program.
For four or more present
categories PROSE could never synthesize a consistent program.

@@

\Cref{fig:recall-precision-categorycount-KWS} shows the same
measurements for an increasing number of structural categories in the
training and test examples.
For more than one structural category in
the training and test examples the recall decreases by about 20\% and
the precision decreases about 6\%.
For more than two structural
categories there is no clear trend in precision, recall, or
F$_{1}$-score.

@@

\noindent
\textbf{Configuration and Input.}
Our study results show that chunk retrieval with PBE gives best
results when the log chunks are structurally identical.
This is
because PROSE has difficulty synthesizing OR-based
programs~\cite{mayer2015user}.
PBE is thus suited to retrieve information
chunks that always have the same defining surrounding or internal
structure.
To extract for example the reason a build failed, the log
passage describing the failure would always have to start and
end the same way.

@@

% shouldn't it also better be in threats to validity?
% it is as well.
%back for my thesis you said we should put it
% directly here to make it more prominent avoid angry critics
% I think it's good here
% but it's also self-contained so feel free to comment it out
\textbf{Caveat!}
This is a preliminary hypothesis based on the results
from our comparison study.
The recommendations could therefore be
influenced by idiosyncrasies of our specific implementation of the
chunk retrieval techniques as well as the \emph{LogChunks}
data set.

@@

\textbf{Implementation.}
Our results depend on the implementation of the investigated chunk
retrieval techniques and the libraries we used.
The program synthesis provided by PROSE is the basis for our
implementation of PBE\@.
The
idiosyncrasies of this framework influence the PBE results.
Other
frameworks similar to PROSE might have somewhat different strengths
and weaknesses.
For example, the need for examples from a single
structural category stems from the fact that PROSE cannot learn
regular expression programs with arbitrary boolean conjunctions and
disjunctions~\cite{mayer2015user}.
PROSE introduced this constraint to
keep the synthesis performance reasonable.
At the same time, this is
clearly a current theoretical challenge of all PBE implementations,
and we therefore attribute it to the technique itself, rather than the
specific implementation.

The library
{\tt text2vec}~\cite{text2vec2019webpage}
and the way it splits strings into word tokens is influencing
our implementation of CTS\@.
On the other hand,
{\tt text2vec} is the de-facto standard library to do word embeddings
in R.
We intentionally chose a simple, minimally configured or tuned
approach to compare against.
Tuning the text similarity
meta-parameters more to the specific use case of chunk retrieval from
build logs would yield better chunk retrieval results.

@@

\textbf{Data Set.}
The build logs from the \emph{LogChunks} data set highly affect
the outcomes of the comparison study.
\emph{LogChunks} consists of build
logs from open source projects and therefore it is not clear whether
our results apply to industry projects that use build tools
not popular in open source development.
However, since \emph{LogChunks} covers a wide array of build tools and
programming languages,
we believe that the findings do generalize.
We collected build logs from Travis CI.
Yet, the format of the log chunk we chose
for the evaluation is largely independent of Travis CI\@.
This
is because the reason the build failed is described within the build
logs by the tools themselves and not the Travis CI environment.

\textbf{Training Set Size.}
Especially the results for CTS might be influenced by the fact that we
only trained on one to five examples.
We chose this small training set
size as a user has to provide the training examples
per repository or project.
We expect a developer or a researcher
to not want to provide more examples than the
small numbers we evaluated on.
% TODO maybe reference back to little effort technique in lit discus.
The fact that PBE and KWS performed best with already two training
examples shows that the training set size is large enough to
compare the three techniques.

@@

\textbf{Usability Analysis of Chunk Retrieval Output.}
Our
analysis of the chunk retrieval output focuses on
precision and recall.
A next step could be to investigate how useful these
outputs really are to developers through controlled experiments or
user studies.
